---
title: "wk5hw"
author: "listianingrumr"
date: "2022-11-14"
output:
  html_document: default
  pdf_document: default
---


# Original week4 homework by Eris (with minor changes)
## load the data first
Here I simplify the file path because you already load the here package.

I also mute the Annex data because it is not needed.
```{r}
library(here)
#Globalgenderinequality <- read.csv("HDR21-22_Statistical_Annex_GII_Table.csv")
GIIdata <- read.csv("Composite_indices_complete_time_series.csv")
library(sf)
Worldshape <- st_read("World_Countries_(Generalized)/World_Countries__Generalized_.shp")
```
## select the data from 2010 and 2019
```{r}
library(dplyr)
GIIdata_2010_19<- GIIdata %>%
  dplyr::select (contains("gii_2010"),
                contains("gii_2019"),
                contains("country"))
```
## creating new column
I also select gii_2010 and gii_2019 because we need the index pop up to make the advanced interactive map.
```{r}
Different_2010_19 <-GIIdata_2010_19 %>%
  mutate(difference=(gii_2010 - gii_2019))%>%
  dplyr::select(gii_2010,gii_2019,difference,
                country)%>%
  
  arrange(desc(difference))
```
## load the countrycode package and add country code table
Before we join the data, we have to add the countrycode because the shapefile use the different country name.
```{r}
library(countrycode)
Different_2010_19$iso = countrycode(Different_2010_19$country,"country.name","iso2c")
```
notes: fyi because you choose country name as the left join id, this warning appears:

Warning: Some values were not matched unambiguously: Arab States, East Asia and the Pacific, Europe and Central Asia, High human development, Latin America and the Caribbean, Low human development, Medium human development, South Asia, Sub-Saharan Africa, Very high human development, World.

It appears because the "naming" system in the csv file is different from the packages, so it cannot add code for some of the countries' names from the list.

## join the data 
```{r}
library(janitor)
library(stringr)
giifinalmap <- Worldshape %>% 
  clean_names() %>% 
  left_join(., Different_2010_19, by = c("country" = "country"))
```
## plot
```{r}
library(tmap)
library(tmaptools)
  tm_shape(giifinalmap) + 
  tm_polygons("difference", 
              style="pretty",
              palette="Blues",
              midpoint=NA,
              #title="Gender Inequality Difference 2010 & 2019",
              alpha = 0.7) + 
  tm_compass(position = c("left", "bottom"),type = "arrow") + 
  tm_scale_bar(position = c("left", "bottom")) +
  tm_layout(title = "Gender Inequality Difference 2010 & 2019", legend.position = c("right", "bottom")) +
  tmap_options(max.categories = 115)
```

# Week5 homework: Map making
## static map
```{r}
tmap_mode("plot")

# set the breaks
# for our mapped data
breaks = c(-0.1, 0, 0.1, 0.2, 0.3, 0.4) 

# plot each map
tm1 <- tm_shape(giifinalmap) + 
  tm_polygons("difference", 
              breaks=breaks,
              palette="PuBu")+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("(a)", position=c(0,0.85), size=1)

legend <- tm_shape(giifinalmap) +
    tm_polygons("difference",
                palette="PuBu") +
    tm_scale_bar(position=c(0.2,0.04), text.size=0.6)+
    tm_compass(north=0, position=c(0.65,0.6))+
    tm_layout(legend.only = TRUE, legend.position=c(0.2,0.25),asp=0.1)+
    tm_credits("(b) difference in gender inequality ", 
               position=c(0.2,0.0), size=0.6)
  
t=tmap_arrange(tm1, legend, ncol=1)

t
```

## basic interactive map
```{r}
tmap_mode("view")

tm_shape(giifinalmap) + 
  tm_polygons("difference", breaks=breaks) 
```
## advanced interactive map
```{r}
# library for pop up boxes (setting 3 pop up boxes:gii_2010, gii_2019 and difference)
library(leafpop)
library(leaflet)

#avoid the geometry for our pop up boxes
popup_gii2010 <-giifinalmap %>%
  st_drop_geometry()%>%
  dplyr::select(`gii_2010`, country)%>%
  popupTable()

popup_gii2019 <-giifinalmap %>%
  st_drop_geometry()%>%
  dplyr::select(`gii_2019`, country)%>%
  popupTable()

popup_difference <-giifinalmap %>%
  st_drop_geometry()%>%
  dplyr::select(`difference`, country)%>%
  popupTable()

tmap_mode("view")

## set the colour palettes using our previous breaks

breaks2=c(0,0.3,0.5,0.7,0.9,1)

pal1 <- giifinalmap %>%
  colorBin(palette = "YlOrRd", domain=.$`gii_2010`, bins=breaks2)

pal2 <- giifinalmap %>%
  colorBin(palette = "YlOrRd", domain=.$`gii_2019`, bins=breaks2)

pal3 <- colorBin(palette = "YlOrRd", 
                domain=giifinalmap$`difference`, bins=breaks)

map<- leaflet(giifinalmap) %>%
  # add basemap options
  addTiles(group = "OSM (default)") %>%
  addProviderTiles(providers$Stamen.Toner, group = "Toner") %>%
  addProviderTiles(providers$Stamen.TonerLite, group = "Toner Lite") %>%
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB")%>%

  #add our polygons, linking to the tables we just made
  addPolygons(color="white", 
              weight = 2,
              opacity = 1,
              dashArray = "3",
              popup = popup_gii2010,
              fillOpacity = 0.7,
              fillColor = ~pal1(`gii_2010`),
              group = "gii_2010") %>%
  
  addPolygons(fillColor = ~pal2(`gii_2019`), 
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              popup = popup_gii2019,
              fillOpacity = 0.7,
              group = "gii_2019") %>%

  addPolygons(color="white", 
              weight = 2,
              opacity = 1,
              dashArray = "3",
              popup = popup_difference,
              fillOpacity = 0.7,
              fillColor = ~pal3(`difference`),
              group = "difference") %>%

  # add a legend

  addLegend(pal = pal3, values = ~`difference`, 
            group = c("gii_2010","gii_2019","difference"), 
            position ="bottomleft", title = "gender inequality index") %>%
  # specify layers control
  addLayersControl(
    baseGroups = c("OSM (default)", "Toner", "Toner Lite", "CartoDB"),
    overlayGroups = c("gii_2010", "gii_2019","difference"),
    options = layersControlOptions(collapsed = FALSE)
  )
## plot the map
map
```